---
title: "Racial Bias Analysis"
output: 
  html_document:
    keep_md: yes
---

# 1. Load Packages
```{r}
library(tidyverse)
library(broom)
library(snakecase)
library(stargazer)
library(gtsummary)
library(broom.helpers)
library(gt)
```

# 2. Load data - Remove invalid trials - Trials with unaffiliated pedestrians
```{r}
data_file <- read_csv("master_file.csv") %>% 
  mutate(
    ethnicity = as.factor(ethnicity),
    gender = as.factor(gender),
    location = as.factor(location),
    first_car_yield = as.factor(first_car_yield),
    did_car_proceed_before_across = as.factor(did_car_proceed_before_across)
  )

data <- data_file %>% drop_na(valid_trial)

data_19th_rm <- data %>% filter(location != "19th")
```

# 3. General descriptive statistics
```{r}
data %>% 
  summarise(
    n = n(),
    mean_time = mean(time_to_cross_street, na.rm = TRUE),
    sd_time = sd(time_to_cross_street, na.rm = TRUE), 
    min_time = min(time_to_cross_street, na.rm = TRUE),
    max_time = max(time_to_cross_street, na.rm = TRUE),
    mean_cars = mean(num_cars_pass_before_yield, na.rm = TRUE), 
    sd_cars = sd(num_cars_pass_before_yield, na.rm = TRUE)
  )
```

# 4. Frequency tables 
```{r}
data %>% count(ethnicity)
data %>% count(gender)
data %>% count(location)
data %>% count(first_car_yield)
data %>% count(did_car_proceed_before_across)
```
# 5. First car yield
# 5a. First car yield - Descriptive stats
```{r}
data %>% 
  group_by(gender) %>% 
  count(first_car_yield) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"

data %>% 
  group_by(ethnicity) %>% 
  count(first_car_yield) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
  
data %>% 
  group_by(ethnicity, gender) %>% 
  count(first_car_yield) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
```

# 5b. First car yield - Logistic regression 
# 5b1. Gender - Location fixed effects
```{r}
m1 <- glm(first_car_yield ~ gender + factor(location),
          data = data,
          family = binomial())

tidy(m1)

summary(m1)

exp(cbind(OR = coef(m1), confint(m1)))
ci1 <- exp(confint(m1))
```

# 5b2. Ethnicity - Location fixed effects
```{r}
m2 <- glm(first_car_yield ~ ethnicity + factor(location),
          data = data,
          family = binomial())

tidy(m2)

summary(m2)

exp(cbind(OR = coef(m2), confint(m2)))
ci2 <- exp(confint(m2))
```

# 5b3. Gender + Ethnicity - Location fixed effects
```{r}
m3 <- glm(first_car_yield ~ gender + ethnicity + factor(location),
          data = data,
          family = binomial())

tidy(m3)

summary(m3)

exp(cbind(OR = coef(m3), confint(m3)))
ci3 <- exp(confint(m3))
```

# 5b4. Gender*Ethnicity - Location fixed effects
```{r}
m4 <- glm(first_car_yield ~ gender*ethnicity + factor(location),
          data = data,
          family = binomial())

tidy(m4)

summary(m4)

exp(cbind(OR = coef(m4), confint(m4)))
ci4 <- exp(confint(m4))
```
# 5c1. Combined models - Odd Ratio with 95% CI - Stargazer
```{r}
stargazer(m1, m2, m3, m4,
          coef = list(exp(coef(m1)),
                      exp(coef(m2)),
                      exp(coef(m3)),
                      exp(coef(m4))),
          ci = TRUE, 
          ci.custom = list(ci1, ci2, ci3, ci4),
          type = "text",
          title = "First Car Yield Models",
          column.labels = c("Gender Model",
                            "Ethnicity Model",
                            "Gender + Ethnicity Model",
                            "Gender*Ethnicity Model"),
          covariate.labels = c("Woman (vs Man)",
                               "White (vs South Asian)",
                               "2nd Avenue",
                               "Bessborough",
                               "Victoria Avenue",
                               "White Woman"),
          dep.var.labels = c("First Car Yield"))

```

# 5c2. Combined models - Odd Ratio with 95% CI - gtsummary
```{r}
tbl1 <- tbl_regression(
  m1,
  exponentiate = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl2 <- tbl_regression(
  m2,
  exponentiate = TRUE,
  label = list(
    ethnicity ~ "Racialization",
    "factor(location)" ~ "Intersection Location"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl3 <- tbl_regression(
  m3,
  exponentiate = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location",
    ethnicity ~ "Racialization"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl4 <- tbl_regression(
  m4,
  exponentiate = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location",
    ethnicity ~ "Racialization"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          label == "woman * white" ~ "Women * White",
          TRUE ~ label)
    ))
```

# 5c3. Table 1 - First car yield combined models table
```{r}
table1 <- tbl_merge(
  tbls = list(tbl1, tbl2, tbl3, tbl4),
  tab_spanner = c(
    "Gender",
    "Racialization",
    "Gender and Racialization",
    "Gender and Racialization Interaction"
  )
) %>% 
  modify_table_body(
    ~ .x %>% 
      mutate(
        row_order = case_when(
          variable == "gender" ~ 1,
          variable == "ethnicity" ~ 2,
          variable == "gender:ethnicity" ~ 3,
          variable == "location" ~ 4,
          TRUE ~ 99
        )
      ) %>% 
      arrange(row_order, row_type != "label") %>% 
      select(-row_order)
  )

as_gt(table1) %>% 
  gtsave(
    filename = "table1.png",
    vwidth = 2200,
    zoom = 2
  )
```

# 6. Mean number of cars before yield
# 6a. Mean number of cars before yield - Descriptive stats
```{r}
data %>% 
  group_by(gender) %>% 
  summarise(
    n = n(),
    mean = mean(num_cars_pass_before_yield),
    sd = sd(num_cars_pass_before_yield),
    .groups = "drop"
  )

data %>% 
  group_by(ethnicity) %>% 
  summarise(
    n = n(),
    mean = mean(num_cars_pass_before_yield),
    sd = sd(num_cars_pass_before_yield),
    .groups = "drop"
  )

data %>% 
  group_by(ethnicity, gender) %>% 
  summarise(
    n = n(),
    mean = mean(num_cars_pass_before_yield),
    sd = sd(num_cars_pass_before_yield),
    .groups = "drop"
  )
```

# 6b. Mean number of cars before yield - Linear regression 
# 6b1. Gender - Location fixed effects
```{r}
m5 <- lm(num_cars_pass_before_yield ~ gender + factor(location),
         data = data)

tidy(m5)

summary(m5)
```

# 6b2. Ethnicity - Location fixed effects
```{r}
m6 <- lm(num_cars_pass_before_yield ~ ethnicity + factor(location),
         data = data)

tidy(m6)

summary(m6)
```

# 6b3. Gender + Ethnicity - Location fixed effects
```{r}
m7 <- lm(num_cars_pass_before_yield ~ gender + ethnicity + factor(location),
         data = data)

tidy(m7)

summary(m7)
```

# 6b4. Gender*Ethnicity - Location fixed effects
```{r}
m8 <- lm(num_cars_pass_before_yield ~ gender*ethnicity + factor(location),
         data = data)

tidy(m8)

summary(m8)
```

# 6c1. Combined models - Stargazer
```{r}
stargazer(m5, m6, m7, m8,
          type = "text",
          title = "Mean Number of Cars that Passed",
           column.labels = c("Gender Model",
                            "Ethnicity Model",
                            "Gender + Ethnicity Model",
                            "Gender*Ethnicity Model"),
          covariate.labels = c("Woman (vs Man)",
                               "White (vs South Asian)",
                               "2nd Avenue",
                               "Bessborough",
                               "Victoria Avenue",
                               "White Woman"),
          dep.var.labels = c("Mean Number of Cars that Passed"))
```

# 6c2. Combined models - gtsummary
```{r}
tbl5 <- tbl_regression(
  m5,
  intercept = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location"
    )
  )%>%
  modify_column_unhide(columns = std.error) %>% 
  modify_column_hide(columns = conf.low) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "(Intercept)" ~ "Constant",
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl6 <- tbl_regression(
  m6,
  intercept = TRUE,
  label = list(
    ethnicity ~ "Racialization",
    "factor(location)" ~ "Intersection Location"
  )) %>% 
  modify_column_unhide(columns = std.error) %>% 
  modify_column_hide(columns = conf.low) %>%
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "(Intercept)" ~ "Constant",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl7 <- tbl_regression(
  m7,
  intercept = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location",
    ethnicity ~ "Racialization"
  )) %>% 
  modify_column_unhide(columns = std.error) %>% 
  modify_column_hide(columns = conf.low) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "(Intercept)" ~ "Constant",
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl8 <- tbl_regression(
  m8,
  intercept = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location",
    ethnicity ~ "Racialization"
  )) %>% 
  modify_column_unhide(columns = std.error) %>% 
  modify_column_hide(columns = conf.low) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "(Intercept)" ~ "Constant",
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          label == "woman * white" ~ "Women * White",
          TRUE ~ label)
    ))
```

# 6c3. Table 2 - Mean number of cars before yield combined models table
```{r}
table2 <- tbl_merge(
  tbls = list(tbl5, tbl6, tbl7, tbl8),
  tab_spanner = c(
    "Gender",
    "Racialization",
    "Gender and Racialization",
    "Gender and Racialization Interaction"
  )
) %>%
  modify_table_body(
    ~ .x %>% 
      mutate(
        row_order = case_when(
          variable == "(Intercept)" ~ 1,
          variable == "gender" ~ 2,
          variable == "ethnicity" ~ 3,
          variable == "gender:ethnicity" ~ 4,
          variable == "location" ~ 5,
          TRUE ~ 99
        )
      ) %>% 
      arrange(row_order, row_type != "label") %>% 
      select(-row_order)
  ) %>% 
  remove_abbreviation("CI = Confidence Interval")

as_gt(table2) %>% 
  gtsave(
    filename = "table2.png",
    vwidth = 2200,
    zoom = 2
  )
```

# 7. Mean time to enter intersection
# 7a. Mean time to enter intersection - Descriptive stats
```{r}
data %>% 
  group_by(gender) %>% 
  summarise(
    n = n(),
    mean = mean(time_to_cross_street),
    sd = sd(time_to_cross_street),
    .groups = "drop"
  )

data %>% 
  group_by(ethnicity) %>% 
  summarise(
    n = n(),
    mean = mean(time_to_cross_street),
    sd = sd(time_to_cross_street),
    .groups = "drop"
  )

data %>% 
  group_by(gender, ethnicity) %>% 
  summarise(
    n = n(),
    mean = mean(time_to_cross_street),
    sd = sd(time_to_cross_street),
    .groups = "drop"
  )
```

# 7b. Mean time to enter intersection - Linear regression
# 7b1. Gender - Location fixed effects
```{r}
m9 <- lm(time_to_cross_street ~ gender + factor(location),
         data = data)

tidy(m9)

summary(m9)
```

# 7b2. Ethnicity - Location fixed effects
```{r}
m10 <- lm(time_to_cross_street ~ ethnicity + factor(location),
         data = data)

tidy(m10)

summary(m10)
```

# 7b3. Gender + Ethnicity - Location fixed effects
```{r}
m11 <- lm(time_to_cross_street ~ gender + ethnicity + factor(location),
         data = data)

tidy(m11)

summary(m11)
```

# 7b4. Gender*Ethnicity - Location fixed effects
```{r}
m12 <- lm(time_to_cross_street ~ gender*ethnicity + factor(location),
         data = data)

tidy(m12)

summary(m12)
```

# 7c1. Combined models - Stargazer 
```{r}
stargazer(m9, m10, m11, m12,
          type = "text",
          title = "Mean Time to Enter the Intersection",
           column.labels = c("Gender Model",
                            "Ethnicity Model",
                            "Gender + Ethnicity Model",
                            "Gender*Ethnicity Model"),
          covariate.labels = c("Woman (vs Man)",
                               "White (vs South Asian)",
                               "2nd Avenue",
                               "Bessborough",
                               "Victoria Avenue",
                               "White Woman"),
          dep.var.labels = c("Mean Time to Enter the Intersection"))
```
# 7c2. Combined models - gtsummary
```{r}
tbl9 <- tbl_regression(
  m9,
  intercept = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location"
    )
  )%>%
  modify_column_unhide(columns = std.error) %>% 
  modify_column_hide(columns = conf.low) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "(Intercept)" ~ "Constant",
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl10 <- tbl_regression(
  m10,
  intercept = TRUE,
  label = list(
    ethnicity ~ "Racialization",
    "factor(location)" ~ "Intersection Location"
  )) %>% 
  modify_column_unhide(columns = std.error) %>% 
  modify_column_hide(columns = conf.low) %>%
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "(Intercept)" ~ "Constant",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl11 <- tbl_regression(
  m11,
  intercept = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location",
    ethnicity ~ "Racialization"
  )) %>% 
  modify_column_unhide(columns = std.error) %>% 
  modify_column_hide(columns = conf.low) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "(Intercept)" ~ "Constant",
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl12 <- tbl_regression(
  m12,
  intercept = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location",
    ethnicity ~ "Racialization"
  )) %>% 
  modify_column_unhide(columns = std.error) %>% 
  modify_column_hide(columns = conf.low) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "(Intercept)" ~ "Constant",
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          label == "woman * white" ~ "Women * White",
          TRUE ~ label)
    ))
```

# 7c3. Table 3 - Mean time before entering intersection combined models table
```{r}
table3 <- tbl_merge(
  tbls = list(tbl9, tbl10, tbl11, tbl12),
  tab_spanner = c(
    "Gender",
    "Racialization",
    "Gender and Racialization",
    "Gender and Racialization Interaction"
  )
) %>%
  modify_table_body(
    ~ .x %>% 
      mutate(
        row_order = case_when(
          variable == "(Intercept)" ~ 1,
          variable == "gender" ~ 2,
          variable == "ethnicity" ~ 3,
          variable == "gender:ethnicity" ~ 4,
          variable == "location" ~ 5,
          TRUE ~ 99
        )
      ) %>% 
      arrange(row_order, row_type != "label") %>% 
      select(-row_order)
  ) %>% 
  remove_abbreviation("CI = Confidence Interval")

as_gt(table3) %>% 
  gtsave(
    filename = "table3.png",
    vwidth = 2200,
    zoom = 2
  )
```

# 8. Car proceed through intersection
# 8a. Car proceed through intersection - Descriptive stats
```{r}
data %>% 
  group_by(gender) %>% 
  count(did_car_proceed_before_across) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"

data %>% 
  group_by(ethnicity) %>% 
  count(did_car_proceed_before_across) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
  
data %>% 
  group_by(ethnicity, gender) %>% 
  count(did_car_proceed_before_across) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
```

# 8b. Car proceed through intersection - Logistic regression  
# 8b1. Gender - Location fixed effects
```{r}
m13 <- glm(did_car_proceed_before_across ~ gender + factor(location),
           data = data,
           family = binomial()
           )

tidy(m13)

summary(m13)

exp(cbind(OR = coef(m13), confint(m13)))
ci13 <- exp(confint(m13))
```

# 8b2. Ethnicity - Location fixed effects
```{r}
m14 <- glm(did_car_proceed_before_across ~ ethnicity + factor(location),
           data = data,
           family = binomial()
           )

tidy(m14)

summary(m14)

exp(cbind(OR = coef(m14), confint(m14)))
ci14 <- exp(confint(m14))
```

# 8b3. Gender + Ethnicity - Location fixed effects
```{r}
m15 <- glm(did_car_proceed_before_across ~ gender + ethnicity + factor(location),
           data = data,
           family = binomial()
           )

tidy(m15)

summary(m15)

exp(cbind(OR = coef(m15), confint(m15)))
ci15 <- exp(confint(m15))
```

# 8b4. Gender*Ethnicity - Location fixed effects
```{r}
m16 <- glm(did_car_proceed_before_across ~ gender*ethnicity + factor(location),
           data = data,
           family = binomial()
           )

tidy(m16)

summary(m16)

exp(cbind(OR = coef(m16), confint(m16)))
ci16 <- exp(confint(m16))
```

# 8c1. Combined models - Odd Ratio with 95% CI - Stargazer
```{r}
stargazer(m13, m14, m15, m16,
          coef = list(exp(coef(m13)),
                      exp(coef(m14)),
                      exp(coef(m15)),
                      exp(coef(m16))),
          ci = TRUE, 
          ci.custom = list(ci13, ci14, ci15, ci16),
          type = "text",
          title = "Car Proceeded through Intersection",
          column.labels = c("Gender Model",
                            "Ethnicity Model",
                            "Gender + Ethnicity Model",
                            "Gender*Ethnicity Model"),
          covariate.labels = c("Woman (vs Man)",
                               "White (vs South Asian)",
                               "2nd Avenue",
                               "Bessborough",
                               "Victoria Avenue",
                               "White Woman"),
          dep.var.labels = c("Car Proceeded through Intersection"))

```

# 8c2. Combined models - Odd Ratio with 95% CI - gtsummary
```{r}
tbl13 <- tbl_regression(
  m13,
  exponentiate = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl14 <- tbl_regression(
  m14,
  exponentiate = TRUE,
  label = list(
    ethnicity ~ "Racialization",
    "factor(location)" ~ "Intersection Location"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl15 <- tbl_regression(
  m15,
  exponentiate = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location",
    ethnicity ~ "Racialization"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl16 <- tbl_regression(
  m16,
  exponentiate = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location",
    ethnicity ~ "Racialization"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          label == "woman * white" ~ "Women * White",
          TRUE ~ label)
    ))
```

# 8c3. Table 4 - Car Proceed combined models table
```{r}
table4 <- tbl_merge(
  tbls = list(tbl13, tbl14, tbl15, tbl16),
  tab_spanner = c(
    "Gender",
    "Racialization",
    "Gender and Racialization",
    "Gender and Racialization Interaction"
  )
) %>% 
  modify_table_body(
    ~ .x %>% 
      mutate(
        row_order = case_when(
          variable == "gender" ~ 1,
          variable == "ethnicity" ~ 2,
          variable == "gender:ethnicity" ~ 3,
          variable == "location" ~ 4,
          TRUE ~ 99
        )
      ) %>% 
      arrange(row_order, row_type != "label") %>% 
      select(-row_order)
  )

as_gt(table4) %>% 
  gtsave(
    filename = "table4.png",
    vwidth = 2200,
    zoom = 2
  )
```

# 9. Cars stop close or far 
# 9a. Cars stop close or far binning
```{r}
data$car_stop_close_or_far_bin <- ifelse(data$car_stop_close_or_far == "far", 1,
                                  ifelse(data$car_stop_close_or_far == "close", 0, NA))
```

# 9b. Cars stop close or far - Descritpive stats
```{r}
data %>% 
  group_by(gender) %>% 
  count(car_stop_close_or_far) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"

data %>% 
  group_by(ethnicity) %>% 
  count(car_stop_close_or_far) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
  
data %>% 
  group_by(ethnicity, gender) %>% 
  count(car_stop_close_or_far) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
```

# 9c. Cars stop close or far - Logistic regression 
# 9c1. Gender - Location fixed effects
```{r}
m17 <- glm(car_stop_close_or_far_bin ~ gender + factor(location),
           data = data,
           family = binomial()
)

tidy(m17)

summary(m17)

exp(cbind(OR = coef(m17), confint(m17)))
ci17 <- exp(confint(m17))
```

# 9c2. Ethnicity - Location fixed effects
```{r}
m18 <- glm(car_stop_close_or_far_bin ~ ethnicity + factor(location),
           data = data,
           family = binomial()
           )

tidy(m18)

summary(m18)

exp(cbind(OR = coef(m18), confint(m18)))
ci18 <- exp(confint(m18))
```

# 9c3. Gender + Ethnicity - Location fixed effects
```{r}
m19 <- glm(car_stop_close_or_far_bin ~ gender + ethnicity + factor(location),
           data = data,
           family = binomial()
           )

tidy(m19)

summary(m19)

exp(cbind(OR = coef(m19), confint(m19)))
ci19 <- exp(confint(m19))
```

# 9c4. Gender*Ethnicity - Location fixed effects
```{r}
m20 <- glm(car_stop_close_or_far_bin ~ gender*ethnicity + factor(location),
           data = data,
           family = binomial()
           )

tidy(m20)

summary(m20)

exp(cbind(OR = coef(m20), confint(m20)))
ci20 <- exp(confint(m20))
```

# 9c1. Combined models - Odd Ratio with 95% CI - Stargazer
```{r}
stargazer(m17, m18, m19, m20,
          coef = list(exp(coef(m17)),
                      exp(coef(m18)),
                      exp(coef(m19)),
                      exp(coef(m20))),
          ci = TRUE, 
          ci.custom = list(ci17, ci18, ci19, ci20),
          type = "text",
          title = "Car Stopped Close or Far Model",
          column.labels = c("Gender Model",
                            "Ethnicity Model",
                            "Gender + Ethnicity Model",
                            "Gender*Ethnicity Model"),
          covariate.labels = c("Woman (vs Man)",
                               "White (vs South Asian)",
                               "2nd Avenue",
                               "Bessborough",
                               "Victoria Avenue",
                               "White Woman"),
          dep.var.labels = c("Car Stopped Close or Far"))

```

# 9c2. Combined models - Odd Ratio with 95% CI - gtsummary
```{r}
tbl17 <- tbl_regression(
  m17,
  exponentiate = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl18 <- tbl_regression(
  m18,
  exponentiate = TRUE,
  label = list(
    ethnicity ~ "Racialization",
    "factor(location)" ~ "Intersection Location"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl19 <- tbl_regression(
  m19,
  exponentiate = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location",
    ethnicity ~ "Racialization"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          TRUE ~ label)
    ))

tbl20 <- tbl_regression(
  m20,
  exponentiate = TRUE,
  label = list(
    gender ~ "Gender",
    "factor(location)" ~ "Intersection Location",
    ethnicity ~ "Racialization"
  )) %>% 
  modify_table_body(
    ~.x %>% 
      mutate(
        label = case_when(
          label == "man" ~ "Men",
          label == "woman" ~ "Women",
          label == "asian" ~ "South Asian",
          label == "white" ~ "White",
          label == "black" ~ "Black",
          label == "19th" ~ "19th Street",
          label == "2nd"  ~ "2nd Avenue",
          label == "bessborough" ~ "Bessborough",
          label == "victoria" ~ "Victoria Avenue",
          label == "woman * white" ~ "Women * White",
          TRUE ~ label)
    ))
```

# 9c3. Table 5 - Car yield close//far combined models table
```{r}
table5 <- tbl_merge(
  tbls = list(tbl17, tbl18, tbl19, tbl20),
  tab_spanner = c(
    "Gender",
    "Racialization",
    "Gender and Racialization",
    "Gender and Racialization Interaction"
  )
) %>% 
  modify_table_body(
    ~ .x %>% 
      mutate(
        row_order = case_when(
          variable == "gender" ~ 1,
          variable == "ethnicity" ~ 2,
          variable == "gender:ethnicity" ~ 3,
          variable == "location" ~ 4,
          TRUE ~ 99
        )
      ) %>% 
      arrange(row_order, row_type != "label") %>% 
      select(-row_order)
  )

as_gt(table5) %>% 
  gtsave(
    filename = "table5.png",
    vwidth = 2200,
    zoom = 2
  )
```

# 10. Histograms
# 10a. Ethnicity
```{r}
data %>% 
  ggplot(aes(x = time_to_cross_street,
         fill = ethnicity)) +
  geom_histogram((aes(y = after_stat(density))),binwidth = 0.5) +
  stat_function(fun = dnorm, args = list(mean = mean(data$time_to_cross_street), sd = sd(data$time_to_cross_street)), colour = "red", linewidth = 1) +
    labs(
    x = "Time to Cross (s)",
    y = "Count"
  ) +
  theme_minimal(
  )
```

# 10b. Gender
```{r}
data %>% 
  ggplot(aes(x = time_to_cross_street,
         fill = gender)) +
  geom_histogram((aes(y = after_stat(density))),binwidth = 0.5) +
  stat_function(fun = dnorm, args = list(mean = mean(data$time_to_cross_street), sd = sd(data$time_to_cross_street)), colour = "red", linewidth = 1) +
  labs(
    x = "Time to Cross (s)",
    y = "Count"
  ) +
  theme_minimal(
  )
```

# 11. Visualizations
# 11a. Time by ethnicity
```{r}
data %>% 
  ggplot(aes(x = ethnicity,
         y = time_to_cross_street,
         fill = ethnicity)) +
  geom_boxplot() +
  labs(
    x = "Ethnicicty",
    y = "Time to Cross (s)"
  ) +
  scale_x_discrete(labels = c(
    "white" = "White",
    "asian" = "South Asian",
    "black" = "Black"
   )) +
  theme_minimal(
  )
```

# 11b. Time by gender
```{r}
data %>% 
  ggplot(aes(x = gender,
         y = time_to_cross_street,
         fill = gender)) +
  geom_boxplot(show.legend = FALSE) +
  labs(
    x = "Gender",
    y = "Time to Cross (s)"
  ) + 
  scale_x_discrete(labels = c(
    "man" = "Man",
    "woman" = "Woman"
   )) +
  theme_minimal()
```

# 11c. Time to enter the street - Violin Plot with Jitter overlay - Facet by gender
```{r}
data %>% 
  ggplot(aes(x = time_to_cross_street,
             y = gender)) +
  geom_violin() +
  geom_jitter(show.legend = FALSE) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE, linewidth = 1.2) +
  labs(
    x = "Time to Cross (s)",
    y = "Gender"
  ) + 
    scale_y_discrete(labels = c(
    "man" = "Man",
    "woman" = "Woman"
   )) +
  theme_minimal()
```

# 11d. Time to enter the street - Violin Plot with Jitter overlay - Facet by ethnicity
```{r}
data %>% 
  ggplot(aes(x = time_to_cross_street,
             y = ethnicity)) +
  geom_violin() +
  geom_jitter(show.legend = FALSE) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE, linewidth = 1.2) +
  labs(
    x = "Time to Cross (s)",
    y = "Gender"
  ) + 
    scale_y_discrete(labels = c(
    "asian" = "South Asian",
    "white" = "White",
    "black" = "Black"
   )) +
  theme_minimal()
```

# 11e. Time to enter - QQplot - By ethnicity  
```{r}
data %>% 
  ggplot(aes(sample = time_to_cross_street)) +
  geom_qq() +
  stat_qq_line() +
  facet_wrap(~ethnicity)
```

# 11f. Time to enter - QQplot - By gender 
```{r}
data %>% 
  ggplot(aes(sample = time_to_cross_street)) +
  geom_qq() +
  stat_qq_line() +
  facet_wrap(~gender)
```

# 11g. 19th removed - Time to enter - By gender
```{r}
data_19th_rm %>% 
  ggplot(aes(x = gender,
         y = time_to_cross_street,
         fill = gender)) +
  geom_boxplot(show.legend = FALSE) +
  labs(
    x = "Gender",
    y = "Time to Cross (s)"
  ) + 
  scale_x_discrete(labels = c(
    "man" = "Man",
    "woman" = "Woman"
  )) +
  theme_minimal()
```

# 11h. Time by ethnicity and gender - grouped by location 
```{r}
data%>% 
  ggplot(aes(x = ethnicity,
             y = time_to_cross_street,
               fill = gender)) +
  geom_boxplot() +
  facet_wrap(~location) +
  labs(
    x = "Ethnicity",
    y = "Time to Cross (s)",
    fill = "Gender"
  ) +
  theme_minimal()
```

# 11i. Cars passed by ethnicity and gender
```{r}
data %>% 
  ggplot(aes(x = ethnicity,
             y = num_cars_pass_before_yield,
             fill = gender)) +
  geom_col() +
  labs (
    x = "Ethnicity",
    y = "Number of Cars Passed",
    fill = "Gender"
  ) +
   scale_x_discrete(labels = c(
    "asian" = "South Asian",
    "white" = "White",
    "black" = "Black"
   )) + 
  theme_minimal()
```

# 11j. Proportion of first car yields - By ethnicity 
```{r}
data %>% 
  ggplot(aes(x = ethnicity,
             fill = first_car_yield)
         ) +
  geom_bar(position = "fill") +
  labs (
    x = "Ethnicity",
    y = "Proportion",
    fill = "Did the First Car Yield"
  ) +
   scale_x_discrete(labels = c(
    "asian" = "South Asian",
    "white" = "White",
    "black" = "Black"
   )) + 
  theme_minimal()
```

# 11k. Proportion of first car yield - By gender 
```{r}
data %>% 
  ggplot(aes(x = gender,
             fill = first_car_yield)
         ) +
  geom_bar(position = "fill") +
  labs (
    x = "Gender",
    y = "Proportion",
    fill = "Did the First Car Yield"
  ) +
  scale_x_discrete(labels = c(
    "man" = "Man",
    "woman" = "Woman"
  )) +
  theme_minimal()
```

# 11l. Odds first car yield
```{r}
or1 <- tidy(m4, conf.int = TRUE, exponentiate = TRUE)

ggplot(or1[-1,], aes(x = estimate,
                    y = reorder(term, estimate))) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = conf.low,
                     xmax = conf.high),
                 height = 0.2) +
  geom_vline(xintercept = 1,
             linetype = "dashed",
             colour = "red") +
  scale_x_log10() +
  scale_y_discrete(labels = c(
    "genderman" = "Man (vs Woman)",
    "genderwoman" = "Woman (vs Man)",
    "ethnicityasian" = "South Asian",
    "ethnicityblack" = "Black",
    "ethnicitywhite" = "White",
    "genderwoman:ethnicitywhite" = "White & Woman",
    "factor(location)victoria" = "Victoria",
    "factor(location)bessborough" = "Bessborough",
    "factor(location)2nd" = "2nd",
    "factor(location)19th" = "19th")) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = "",
    title = "Logistic Regression Results"
  ) +
  theme_minimal()
```

# 11m. Odds car proceeded
```{r}
or2 <- tidy(m16, conf.int = TRUE, exponentiate = TRUE)

ggplot(or2[-1,], aes(x = estimate,
                    y = reorder(term, estimate))) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = conf.low,
                     xmax = conf.high),
                 height = 0.2) +
  geom_vline(xintercept = 1,
             linetype = "dashed",
             colour = "red") +
  scale_x_log10() +
  scale_y_discrete(labels = c(
    "genderman" = "Male (vs Female)",
    "genderwoman" = "Female (vs Male)",
    "ethnicityasian" = "South Asian",
    "ethnicityblack" = "Black",
    "ethnicitywhite" = "White",
    "factor(location)victoria" = "Victoria",
    "factor(location)bessborough" = "Bessborough",
    "factor(location)2nd" = "2nd",
    "factor(location)19th" = "19th",
    "genderwoman:ethnicitywhite" = "White & Woman")) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = "",
    title = "Logistic Regression Results"
  ) +
  theme_minimal()
```

# 11n. Odds car stopped close//far 
```{r}
or3 <- tidy(m20, conf.int = TRUE, exponentiate = TRUE)

ggplot(or3[-1,], aes(x = estimate,
                    y = reorder(term, estimate))) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = conf.low,
                     xmax = conf.high),
                 height = 0.2) +
  geom_vline(xintercept = 1,
             linetype = "dashed",
             colour = "red") +
  scale_x_log10() +
  scale_y_discrete(labels = c(
    "genderman" = "Male (vs Female)",
    "genderwoman" = "Female (vs Male)",
    "ethnicityasian" = "South Asian",
    "ethnicityblack" = "Black",
    "ethnicitywhite" = "White",
    "factor(location)victoria" = "Victoria",
    "factor(location)bessborough" = "Bessborough",
    "factor(location)2nd" = "2nd",
    "factor(location)19th" = "19th",
    "genderwoman:ethnicitywhite" = "White & Woman")) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = "",
    title = "Logistic Regression Results"
  ) +
  theme_minimal()
```