---
title: "Racial Bias Analysis"
output: 
  html_document:
    keep_md: yes
---

LOAD PACKAGES
```{r}
library(tidyverse)
library(broom)
library(snakecase)
```

LOAD DATA - REMOVE INVALID TRIALS - TRIALS WITH UNAFFILIATED PEDESTRIANS
```{r}
data_file <- read_csv("master_file.csv") %>% 
  mutate(
    ethnicity = as.factor(ethnicity),
    gender = as.factor(gender),
    location = as.factor(location),
    first_car_yield = as.factor(first_car_yield),
    did_car_proceed_before_across = as.factor(did_car_proceed_before_across)
  )

data <- data_file %>% drop_na(valid_trial)

data_19th_rm <- data %>% filter(location != "19th")
```

GENERAL DESCRIPTIVE STATS
```{r}
data %>% 
  summarise(
    n = n(),
    mean_time = mean(time_to_cross_street, na.rm = TRUE),
    sd_time = sd(time_to_cross_street, na.rm = TRUE), 
    min_time = min(time_to_cross_street, na.rm = TRUE),
    max_time = max(time_to_cross_street, na.rm = TRUE),
    mean_cars = mean(num_cars_pass_before_yield, na.rm = TRUE), 
    sd_cars = sd(num_cars_pass_before_yield, na.rm = TRUE)
  )
```

FREQUENCY TABLES
```{r}
data %>% count(ethnicity)
data %>% count(gender)
data %>% count(location)
data %>% count(first_car_yield)
data %>% count(did_car_proceed_before_across)
```

FIRST CAR YIELD DESCRIPTIVE STATS
```{r}
data %>% 
  group_by(gender) %>% 
  count(first_car_yield) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"

data %>% 
  group_by(ethnicity) %>% 
  count(first_car_yield) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
  
data %>% 
  group_by(ethnicity, gender) %>% 
  count(first_car_yield) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
```

FIRST CAR YIELD - LOGISTIC REGRESSION
GENDER - LOCATION FIXED EFFECTS
```{r}
m1 <- glm(first_car_yield ~ gender + factor(location),
          data = data,
          family = binomial())

tidy(m1)

summary(m1)

exp(coef(m1))
```

ETHNICITY - LOCATION FIXED EFFECTS
```{r}
m2 <- glm(first_car_yield ~ ethnicity + factor(location),
          data = data,
          family = binomial())

tidy(m2)

summary(m2)

exp(coef(m2))
```

GENDER + ETHNICITY - LOCATION FIXED EFFECTS
```{r}
m3 <- glm(first_car_yield ~ gender + ethnicity + factor(location),
          data = data,
          family = binomial())

tidy(m3)

summary(m3)

exp(coef(m3))
```

MEAN NUMBER OF CARS PASS BEFORE YIELD - DESCRIPTIVE STATS
```{r}
data %>% 
  group_by(gender) %>% 
  summarise(
    n = n(),
    mean = mean(num_cars_pass_before_yield),
    sd = sd(num_cars_pass_before_yield),
    .groups = "drop"
  )

data %>% 
  group_by(ethnicity) %>% 
  summarise(
    n = n(),
    mean = mean(num_cars_pass_before_yield),
    sd = sd(num_cars_pass_before_yield),
    .groups = "drop"
  )

data %>% 
  group_by(ethnicity, gender) %>% 
  summarise(
    n = n(),
    mean = mean(num_cars_pass_before_yield),
    sd = sd(num_cars_pass_before_yield),
    .groups = "drop"
  )
```

MEAN NUMBER OF CARS PASS BEFORE YIELD - LINEAR REGRESSION
GENDER - LOCATION FIXED EFFECTS
```{r}
m4 <- lm(num_cars_pass_before_yield ~ gender + factor(location),
         data = data)

tidy(m4)

summary(m4)
```

ETHNICITY - LOCATION FIXED EFFECTS
```{r}
m5 <- lm(num_cars_pass_before_yield ~ ethnicity + factor(location),
         data = data)

tidy(m5)

summary(m5)
```

GENDER + ETHNICITY - LOCATION FIXED EFFECTS
```{r}
m6 <- lm(num_cars_pass_before_yield ~ gender + ethnicity + factor(location),
         data = data)

tidy(m6)

summary(m6)
```

TIME TO ENTER INTERSECTION - DESCRIPTIVE STATS
```{r}
data %>% 
  group_by(gender) %>% 
  summarise(
    n = n(),
    mean = mean(time_to_cross_street),
    sd = sd(time_to_cross_street),
    .groups = "drop"
  )

data %>% 
  group_by(ethnicity) %>% 
  summarise(
    n = n(),
    mean = mean(time_to_cross_street),
    sd = sd(time_to_cross_street),
    .groups = "drop"
  )

data %>% 
  group_by(gender, ethnicity) %>% 
  summarise(
    n = n(),
    mean = mean(time_to_cross_street),
    sd = sd(time_to_cross_street),
    .groups = "drop"
  )
```

TIME TO ENTER INTERSECTION - LINEAR REGRESSION
GENDER - LOCATION FIXED EFFECTS 
```{r}
m7 <- lm(time_to_cross_street ~ gender + factor(location),
         data = data)

tidy(m7)

summary(m7)
```

ETHNICITY - LOCATION FIXED EFFECTS
```{r}
m8 <- lm(time_to_cross_street ~ ethnicity + factor(location),
         data = data)

tidy(m8)

summary(m8)
```

GENDER + ETHNICITY - LOCATION FIXED EFFECTS
```{r}
m9 <- lm(time_to_cross_street ~ gender + ethnicity + factor(location),
         data = data)

tidy(m9)

summary(m9)
```

CAR PROCEED THROUGH INTERSECTION - DESCRIPTIVE STATS
```{r}
data %>% 
  group_by(gender) %>% 
  count(did_car_proceed_before_across) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"

data %>% 
  group_by(ethnicity) %>% 
  count(did_car_proceed_before_across) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
  
data %>% 
  group_by(ethnicity, gender) %>% 
  count(did_car_proceed_before_across) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
```

CAR PROCEED THROUGH INTERSECTION - LOGISTIC REGRESSION 
GENDER - LOCATION FIXED EFFECTS
```{r}
m10 <- glm(did_car_proceed_before_across ~ gender + factor(location),
           data = data,
           family = binomial()
           )

tidy(m10)

summary(m10)

exp(coef(m10))
```

ETHNICITY - LOCATION FIXED EFFECTS
```{r}
m11 <- glm(did_car_proceed_before_across ~ ethnicity + factor(location),
           data = data,
           family = binomial()
           )

tidy(m11)

summary(m11)

exp(coef(m11))
```

GENDER + ETHNICITY - LOCATION FIXED EFFECTS
```{r}
m12 <- glm(did_car_proceed_before_across ~ gender + ethnicity + factor(location),
           data = data,
           family = binomial()
           )

tidy(m12)

summary(m12)

exp(coef(m12))
```

CARS STOP CLOSE OR FAR BINNING
```{r}
data$car_stop_close_or_far_bin <- ifelse(data$car_stop_close_or_far == "far", 1,
                                  ifelse(data$car_stop_close_or_far == "close", 0, NA))
```

CARS STOP CLOSE OR FAR - DESCRIPTIVE STATS
```{r}
data %>% 
  group_by(gender) %>% 
  count(car_stop_close_or_far) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"

data %>% 
  group_by(ethnicity) %>% 
  count(car_stop_close_or_far) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
  
data %>% 
  group_by(ethnicity, gender) %>% 
  count(car_stop_close_or_far) %>% 
  mutate(percentage = n / sum(n) * 100)
  .groups = "drop"
```

CARS STOP CLOSE OR FAR - LOGISTIC REGRESSION
GENDER - LOCATION FIXED EFFECTS
```{r}
m13 <- glm(car_stop_close_or_far_bin ~ gender + factor(location),
           data = data,
           family = binomial()
)

tidy(m13)

summary(m13)

exp(coef(m13))
```

ETHNICITY - LOCATION FIXED EFFECTS
```{r}
m14 <- glm(car_stop_close_or_far_bin ~ ethnicity + factor(location),
           data = data,
           family = binomial()
           )

tidy(m14)

summary(m14)

exp(coef(m14))
```

GENDER + ETHNICITY - LOCATION FIXED EFFECTS 
```{r}
m15 <- glm(car_stop_close_or_far_bin ~ gender + ethnicity + factor(location),
           data = data,
           family = binomial()
           )

tidy(m15)

summary(m15)

exp(coef(m15))
```

VISUALIZATIONS TIME BY RACE
```{r}
data %>% 
  ggplot(aes(x = ethnicity,
         y = time_to_cross_street,
         fill = ethnicity)) +
  geom_boxplot() +
  labs(
    x = "Race",
    y = "Time to Cross (s)"
  ) +
  theme_minimal(
  )
```

TIME BY GENDER
```{r}
data %>% 
  ggplot(aes(x = gender,
         y = time_to_cross_street,
         fill = gender)) +
  geom_boxplot(show.legend = FALSE) +
  labs(
    x = "Gender",
    y = "Time to Cross (s)"
  ) + 
  theme_minimal()
```

```{r}
data %>% 
  ggplot(aes(x = time_to_cross_street,
             y = gender)) +
  geom_violin() +
  geom_jitter(show.legend = FALSE) +
  geom_smooth(aes(group = 1), method = "lm", se = FALSE, size = 1.2) +
  labs(
    x = "Time to Cross (s)",
    y = "Gender"
  ) + 
  theme_minimal()
```
```{r}
data %>% 
  ggplot(aes(sample = time_to_cross_street)) +
  geom_qq() +
  stat_qq_line() +
  facet_wrap(~gender)
```

19TH REMOVED - TIME TO CROSS BY GENDER
```{r}
data_19th_rm %>% 
  ggplot(aes(x = gender,
         y = time_to_cross_street,
         fill = gender)) +
  geom_boxplot(show.legend = FALSE) +
  labs(
    x = "Gender",
    y = "Time to Cross (s)"
  ) + 
  theme_minimal()
```

TIME BY RACE AND GENDER - GROUPED BY LOCATION
```{r}
data%>% 
  ggplot(aes(x = ethnicity,
             y = time_to_cross_street,
               fill = gender)) +
  geom_boxplot() +
  facet_wrap(~location) +
  labs(
    x = "Race",
    y = "Time to Cross (s)",
    fill = "Gender"
  ) +
  theme_minimal()
```

CARS PASSED BY RACE AND GENDER
```{r}
data %>% 
  ggplot(aes(x = ethnicity,
             y = num_cars_pass_before_yield,
             fill = gender)) +
  geom_col() +
  labs (
    x = "Race",
    y = "Number of Cars Passed",
    fill = "Gender"
  ) +
  theme_minimal()
```

PROPORTION OF FIRST CAR YIELD RACE VISUAL
```{r}
data %>% 
  ggplot(aes(x = ethnicity,
             fill = first_car_yield)
         ) +
  geom_bar(position = "fill") +
  labs (
    x = "Race",
    y = "Proportion",
    fill = "Did the First Car Yield"
  ) +
  theme_minimal()
```

PROPORTION OF FIRST CAR YIELD GENDER VISUAL
```{r}
data %>% 
  ggplot(aes(x = gender,
             fill = first_car_yield)
         ) +
  geom_bar(position = "fill") +
  labs (
    x = "Gender",
    y = "Proportion",
    fill = "Did the First Car Yield"
  ) +
  theme_minimal()
```

ODDS FIRST CAR YIELD
```{r}
or <- tidy(m3, conf.int = TRUE, exponentiate = TRUE)

ggplot(or[-1,], aes(x = estimate,
                    y = reorder(term, estimate))) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = conf.low,
                     xmax = conf.high),
                 height = 0.2) +
  geom_vline(xintercept = 1,
             linetype = "dashed",
             colour = "red") +
  scale_x_log10() +
  scale_y_discrete(labels = c(
    "genderman" = "Male (vs Female)",
    "genderwoman" = "Female (vs Male)",
    "ethnicityasian" = "South Asian",
    "ethnicityblack" = "Black",
    "ethnicitywhite" = "White",
    "factor(location)victoria" = "Victoria",
    "factor(location)bessborough" = "Bessborough",
    "factor(location)2nd" = "2nd",
    "factor(location)19th" = "19th")) +
  labs(
    x = "Odds Ratio (95% CI)",
    y = "",
    title = "Logistic Regression Results"
  ) +
  theme_minimal()
```