LOAD PACKAGES

```{r}
library(tidyverse)
library(broom)
library(snakecase)
```

LOAD DATA - REMOVE INVALID TRIALS - TRIALS WITH UNAFFILIATED PEDESTRIANS

```{r}
data_file <- read_csv("master_file.csv") %>% 
  mutate(
    ethnicity = as.factor(ethnicity),
    gender = as.factor(gender),
    location = as.factor(location),
    first_car_yield = as.factor(first_car_yield),
    did_car_proceed_before_across = as.factor(did_car_proceed_before_across)
  )

data <- data_file %>% drop_na(valid_trial)
```

GENERAL DESCRIPTIVE STATS

```{r}
data %>% 
  summarise(
    n = n(),
    mean_time = mean(time_to_cross_street, na.rm = TRUE),
    sd_time = sd(time_to_cross_street, na.rm = TRUE), 
    min_time = min(time_to_cross_street, na.rm = TRUE),
    max_time = max(time_to_cross_street, na.rm = TRUE),
    mean_cars = mean(num_cars_pass_before_yield, na.rm = TRUE), 
    sd_cars = sd(num_cars_pass_before_yield, na.rm = TRUE)
  )
```

FREQUENCY TABLES

```{r}
data %>% count(ethnicity)
data %>% count(gender)
data %>% count(location)
data %>% count(first_car_yield)
data %>% count(did_car_proceed_before_across)
```

TIME TO CROSS BY RACE DESCRIPTIVE STATS

```{r}
data %>% 
  group_by(ethnicity) %>% 
  summarise(
    n = n(),
    mean = mean(time_to_cross_street),
    sd = sd(time_to_cross_street)
  )
```

ONE-WAY ANOVA RACE

```{r}
race_model <- aov(
  time_to_cross_street~ethnicity,
  data = data
)

tidy(race_model)

TukeyHSD(race_model)
```

ONE-WAY ANOVA LOCATION

```{r}
location_model <- aov(
  time_to_cross_street~location,
  data = data
)

tidy(location_model)

TukeyHSD(location_model)
```

TIME TO CROSS BY GENDER DESCRIPTIVE STATS

```{r}
data %>% 
  group_by(gender) %>% 
  summarise(
    n = n(),
    mean = mean(time_to_cross_street),
    sd = sd(time_to_cross_street)
  )
```

T-TEST

```{r}
t.test(
  time_to_cross_street ~ gender,
  data = data
)
```

TIME TO CROSS BY RACE X GENDER

```{r}
data %>% 
  group_by(ethnicity, gender) %>% 
  summarise(
    n = n(),
    mean = mean(time_to_cross_street),
    sd = sd(time_to_cross_street),
    .groups = "drop"
  )
```

TWO-WAY ANOVA

```{r}
time_model <- aov(
  time_to_cross_street ~ ethnicity*gender,
  data = data
)

tidy(time_model)
```

NUMBER OF CARS PASSED BY GENDER AND/OR RACE DESCRIPTIVE STATS

```{r}
data %>% 
  group_by(ethnicity, gender) %>% 
  summarise(
    n = n(),
    mean = mean(num_cars_pass_before_yield),
    sd = sd(num_cars_pass_before_yield),
    .groups = "drop"
  )
```

GENDER COMPARISON

```{r}
t.test(
  num_cars_pass_before_yield ~ gender,
  data = data
)
```

RACE COMPARISON

```{r}
cars_race <- aov(
  num_cars_pass_before_yield ~ ethnicity,
  data = data
)

tidy(cars_race)
```

LOCATION COMPARISON

```{r}
cars_location <- aov(
  num_cars_pass_before_yield ~ location, 
  data = data
)

tidy(cars_location)

TukeyHSD(cars_location)
```

RACE AND GENDER COMPARISON

```{r}
cars_model <- aov(
  num_cars_pass_before_yield ~ ethnicity * gender,
  data = data
)

tidy(cars_model)
```

FIRST CAR YIELD BY RACE AND/OR GENDER RACE COMPARISON

```{r}
data %>% 
  count(ethnicity, first_car_yield)

chisq.test(
  table(data$ethnicity,
        data$first_car_yield)
)
```

LOCATION COMPARISON

```{r}
data %>% 
  count(location, first_car_yield)

chisq.test(
  table(data$location,
        data$first_car_yield)
)
```

GENDER COMPARISON

```{r}
data %>% 
  count(gender, first_car_yield)

chisq.test(
  table(data$gender,
        data$first_car_yield)
)
```

LOGISTIC REGRESSION

```{r}
firstcar_model <- glm(
  first_car_yield ~ ethnicity + gender,
  data = data,
  family = binomial()
)

tidy(firstcar_model)
```

DID CAR PROCEED BY RACE AND/OR GENDER RACE COMPARISON

```{r}
data %>% 
  count(ethnicity, did_car_proceed_before_across)

chisq.test(
  table(data$ethnicity,
        data$did_car_proceed_before_across)
)
```

LOCATION COMPARISON

```{r}
data %>% 
  count(location, did_car_proceed_before_across)

chisq.test(
  table(data$location,
        data$did_car_proceed_before_across)
)
```

GENDER COMPARISON

```{r}
data %>% 
  count(gender, did_car_proceed_before_across)

chisq.test(
  table(data$gender,
        data$did_car_proceed_before_across)
)
```

LOGISTIC REGRESSION

```{r}
proceed_model <- glm(
  did_car_proceed_before_across ~ ethnicity + gender,
  data = data,
  family = binomial()
)

tidy(proceed_model)
```

DID CAR YIELD CLOSE OR FAR BY RACE AND/OR GENDER GENDER COMPARISON

```{r}
data %>% 
  count(gender, car_stop_close_or_far)

chisq.test(
  table(data$gender,
        data$car_stop_close_or_far)
)
```

RACE COMPARISON

```{r}
data %>% 
  count(ethnicity, car_stop_close_or_far)

chisq.test(
  table(data$ethnicity,
        data$car_stop_close_or_far)
)
```

LOCATION COMPARISON

```{r}
data %>% 
  count(location, car_stop_close_or_far)

chisq.test(
  table(data$location,
        data$car_stop_close_or_far)
)
```

LOGISTIC REGRESSION

```{r}
data$car_stop_close_or_far_bin <- ifelse(data$car_stop_close_or_far == "far", 1,
                                  ifelse(data$car_stop_close_or_far == "close", 0, NA))

yield_model <- glm(
  car_stop_close_or_far_bin ~ ethnicity + gender,
  data = data,
  family = binomial()
)

tidy(yield_model)
```

VISUALIZATIONS TIME BY RACE

```{r}
data %>% 
  ggplot(aes(x = ethnicity,
         y = time_to_cross_street,
         fill = ethnicity)) +
  geom_boxplot() +
  labs(
    x = "Race",
    y = "Time to Cross (s)"
  ) +
  theme_minimal(
  )
```

TIME BY GENDER

```{r}
data %>% 
  ggplot(aes(x = gender,
         y = time_to_cross_street,
         fill = gender)) +
  geom_boxplot(show.legend = FALSE) +
  labs(
    x = "Gender",
    y = "Time to Cross (s)"
  ) + 
  theme_minimal()
```

TIME BY RACE AND GENDER - GROUPED BY LOCATION

```{r}
data %>% 
  ggplot(aes(x = ethnicity,
             y = time_to_cross_street,
               fill = gender)) +
  geom_boxplot() +
  facet_wrap(~location) +
  labs(
    x = "Race",
    y = "Time to Cross (s)",
    fill = "Gender"
  ) +
  theme_minimal()
```

CARS PASSED BY RACE AND GENDER

```{r}
data %>% 
  ggplot(aes(x = ethnicity,
             y = num_cars_pass_before_yield,
             fill = gender)) +
  geom_col() +
  labs (
    x = "Race",
    y = "Number of Cars Passed",
    fill = "Gender"
  ) +
  theme_minimal()
```

PROPORTION OF FIRST CAR YIELD RACE VISUAL

```{r}
data %>% 
  ggplot(aes(x = ethnicity,
             fill = first_car_yield)
         ) +
  geom_bar(position = "fill") +
  labs (
    x = "Race",
    y = "Proportion",
    fill = "Did the First Car Yield"
  ) +
  theme_minimal()
```

PROPORTION OF FIRST CAR YIELD GENDER VISUAL

```{r}
data %>% 
  ggplot(aes(x = gender,
             fill = first_car_yield)
         ) +
  geom_bar(position = "fill") +
  labs (
    x = "Gender",
    y = "Proportion",
    fill = "Did the First Car Yield"
  ) +
  theme_minimal()
```
